---
title: "Exercise_1"
author: "Sam Worthy"
date: "2023-11-26"
output: html_notebook
---

```{r}
library(tidymodels)
library(multilevelmod)
tidymodels_prefer()
```

We are using the built-in ChickWeight dataset

```{r}
data(ChickWeight)
# then it is an object in your environment
summary(ChickWeight)
```

### Use rsample to create training and test sets from ChickWeight

Each chick only received one diet so need to split by chicks acrosss diets

```{r}
chick.split = group_initial_split(ChickWeight,Chick)
chick.split
```

```{r}
chicks.training = training(chick.split)
chicks.testing = testing(chick.split)
```


### Fit Linear Model Additive
Fit a linear model to ChickWeight training dataset, modeling weight as a function of diet and time (and their interaction?), using parsnip tools.

```{r}
lm.model <- 
  linear_reg() %>% 
  set_engine("lm")

chicks.lm.fit <- 
  lm.model %>% 
  fit(weight ~ Diet + Time, data = chicks.training)

chicks.lm.fit
tidy(chicks.lm.fit)
```

### Fit Linear Model Interaction

```{r}
lm.model <- 
  linear_reg() %>% 
  set_engine("lm")

chicks.lm.fit.interact <- 
  lm.model %>% 
  fit(weight ~ Diet + Time + Diet*Time, data = chicks.training)

chicks.lm.fit.interact
tidy(chicks.lm.fit.interact)
```

### Linear mixed effect model

This really should be a linear mixed-effect model so see if you can do that instead (e.g. with lme4 or stan at the engine)

```{r}
lmer.model <- 
  linear_reg() %>% 
  set_engine("lmer")

chicks.lmer.fit <- 
  lmer.model %>% 
  fit(weight ~ Diet + Time + (1|Chick), data = chicks.training)

chicks.lmer.fit
```

```{r}
chicks.lmer.fit %>% extract_fit_engine() %>% summary
```

### Predict Weight 
Use your model to predict weight in your test set chicks using parsnip tools

```{r}
chicks.testing.predict = chicks.testing %>%
  bind_cols(predict(chicks.lm.fit, chicks.testing)) %>% 
  # Add 95% prediction intervals to the results:
  bind_cols(predict(chicks.lm.fit, chicks.testing, type = "pred_int"))
summary(chicks.testing.predict)
```
# predictions are the same for chicks with the same diet?

```{r}
chicks.testing.predict.2 = chicks.testing %>%
  bind_cols(predict(chicks.lm.fit.interact, chicks.testing)) %>% 
  # Add 95% prediction intervals to the results:
  bind_cols(predict(chicks.lm.fit, chicks.testing, type = "pred_int"))
summary(chicks.testing.predict.2)
```

### Plot predicted versus observed in your test data set

```{r}
ggplot(chicks.testing.predict, aes(x = weight, y = .pred, color = Diet)) +
  geom_point()
```

```{r}
ggplot(chicks.testing.predict.2, aes(x = weight, y = .pred, color = Diet)) +
  geom_point()
```


