---
title: "Exercise_02-12-24"
author: "Sam Worthy"
date: "2024-02-11"
output:
  html_document:
    keep_md: yes
---

```{r}
library(tidymodels)
tidymodels_prefer()
library(modeldata)
library(ggplot2)
library(workflowsets)
library(ggrepel)
library(tidyposterior)
library(rstanarm)
```

```{r}
data("Chicago")
Chicago
```

set up the weekend variable:
```{r}
Chicago <- Chicago %>%
  mutate(weekend = timeDate::isBizday(timeDate::as.timeDate(date)))
```


## Training and Test

Make an 80/20 train/test split.  Strata with ridership

```{r}
set.seed(010324)
ride_split = initial_split(Chicago, prop = 0.80, strata = ridership)
ride_train = training(ride_split)
ride_test = testing(ride_split)
```

## 3. Workflow Set

```{r}
lm_model <- 
  linear_reg() %>% 
  set_engine("lm")
```

```{r}
climate = list(
  temp_min = ridership ~ temp_min + weather_storm,
  temp = ridership ~ temp + weather_storm,
  temp_max = ridership ~ temp_max + weather_storm,
  temp_change = ridership ~ temp_change + weather_storm,
  precip = ridership ~ percip + weather_storm,
  precip_max = ridership ~ percip_max + weather_storm)
```

```{r}
climate_models = workflow_set(preproc = climate, models = list(lm = lm_model))
climate_models
```

```{r}
climate_models_fit <-
   climate_models %>%
   mutate(fit = map(info, ~ fit(.x$workflow[[1]], ride_train)))
climate_models_fit
```

## 4 Recipes

```{r}
ride_recipe = 
  recipe(ridership ~ ., data = ride_train) %>%
  update_role(date, new_role = "date") %>%
  step_normalize(Austin:weather_storm) %>%
  step_pca(temp_min:weather_storm,threshold = .75, prefix = "weather") %>%
  step_pca(Austin:California, threshold = .75, prefix = "station")
ride_recipe
```

### 4B
Use the recipe from 4A to fit a linear regression of ridership on the new PCs and all remaining predictors (i.e. those not used in making the PCs).  Use the training data.

```{r}
lm_wflow <- 
  workflow() %>% 
  add_model(lm_model) %>%
  add_recipe(ride_recipe)
```

```{r}
lm_fit_recipe = fit(lm_wflow, ride_train)
lm_fit_recipe
```

## New Exercise Material

Use Cross-fold validation and the model selection techniques shown in chapter 12 to compare models with:

1. All PCs and remaining predictors (same as Q4)
2. The PCs + the weekend variable (no sports team data)
3. 1 weather PC, 1 station PC, + the weekend variable
4. 1 weather PC + the weekend variable
5. 1 station PC + the weekend variable
6. the weekend variable only

New simple recipe

```{r}
ride_recipe = 
  recipe(ridership ~ ., data = ride_train) %>%
  update_role(date, new_role = "date") %>%
  update_role(Austin:California, new_role = "station") %>%
  update_role(temp_min:weather_storm, new_role = "weather") %>%
  update_role(Blackhawks_Away:Cubs_Home, new_role = "sports") %>%
  step_normalize(has_role("station"), has_role("weather"))
ride_recipe
```

adding new small recipes based on models

```{r}
# Build each recipe
rec_1 <- ride_recipe %>%
  step_pca(has_role("station"), threshold = .75, prefix = "station_PC", id = "station_pca") %>%
  step_pca(has_role("weather"), threshold = .75, prefix = "weather_PC", id = "weather_pca") %>%
  update_role(has_role("sports"), new_role = "predictor")

rec_2 <- ride_recipe %>%
  step_pca(has_role("station"), threshold = .75, prefix = "station_PC", id = "station_pca") %>%
  step_pca(has_role("weather"), threshold = .75, prefix = "weather_PC", id = "weather_pca")

rec_3 <- ride_recipe %>%
  step_pca(has_role("station"), num_comp = 1, prefix = "station_PC", id = "station_pca") %>%
  step_pca(has_role("weather"), num_comp = 1, prefix = "weather_PC", id = "weather_pca")
  
rec_4 <- ride_recipe %>%
  step_pca(has_role("weather"), num_comp = 1, prefix = "weather_PC", id = "weather_pca")

rec_5 <- ride_recipe %>%
  step_pca(has_role("station"), num_comp = 1, prefix = "station_PC", id = "station_pca")

rec_6 <- ride_recipe

preproc <- list(
  one = rec_1,
  two = rec_2,
  three = rec_3,
  four = rec_4,
  five = rec_5,
  six = rec_6
)
```


```{r}
lm_models <- workflow_set(preproc, list(lm = linear_reg()), cross = FALSE)
lm_models
```

Get folds

```{r}
chicago_folds <- vfold_cv(ride_train, v = 10)
chicago_folds
```
```{r}
keep_pred <- control_resamples(save_pred = TRUE, save_workflow = TRUE)
```


```{r}
lm_models <- 
  lm_models %>% 
  workflow_map("fit_resamples", 
               seed = 1101, verbose = TRUE,
               # Options to `fit_resamples()`: 
               resamples = chicago_folds, control = keep_pred)
```

```{r}
lm_models
```

```{r}
collect_metrics(lm_models) %>% 
  filter(.metric == "rmse")
```

```{r}
autoplot(lm_models, metric = "rsq") +
  geom_text_repel(aes(label = wflow_id), nudge_x = 1/8, nudge_y = 1/100) +
  theme(legend.position = "none")
```

```{r}
rsq_indiv_estimates <- 
  collect_metrics(lm_models, summarize = FALSE) %>% 
  filter(.metric == "rsq") 

rsq_wider <- 
  rsq_indiv_estimates %>% 
  select(wflow_id, .estimate, id) %>% 
  pivot_wider(id_cols = "id", names_from = "wflow_id", values_from = ".estimate")

corrr::correlate(rsq_wider %>% select(-id), quiet = TRUE)
```

```{r}
rsq_indiv_estimates %>% 
  mutate(wflow_id = reorder(wflow_id, .estimate)) %>% 
  ggplot(aes(x = wflow_id, y = .estimate, group = id, color = id)) + 
  geom_line(alpha = .5, linewidth = 1.25) + 
  theme(legend.position = "none")
```

```{r}
rsq_anova <-
  perf_mod(
    lm_models,
    metric = "rsq",
    prior_intercept = rstanarm::student_t(df = 1),
    chains = 4,
    iter = 5000,
    seed = 1102
  )
```

```{r}
model_post <- 
  rsq_anova %>% 
  # Take a random sample from the posterior distribution
  # so set the seed again to be reproducible. 
  tidy(seed = 1103) 

glimpse(model_post)
```

```{r}
model_post %>% 
  mutate(model = forcats::fct_inorder(model)) %>%
  ggplot(aes(x = posterior)) + 
  geom_histogram(bins = 50, color = "white", fill = "blue", alpha = 0.4) + 
  facet_wrap(~ model, ncol = 1)
```

```{r}
autoplot(rsq_anova) +
  geom_text_repel(aes(label = workflow), nudge_x = 1/8, nudge_y = 1/100) +
  theme(legend.position = "none")
```

```{r}
autoplot(rsq_anova, type = "ROPE", size = 0.02) +
  geom_text_repel(aes(label = workflow)) +
  theme(legend.position = "none")
```

